Kalman tracking and estimation of continuous gravitational waves with a pulsar timing array

### 2023-08-30 docs/manuscript/paper.tex
commit before updating posteriori plots

\subsubsection{Multiple noise realisations} \label{sec:multiple_noise}
\begin{figure*}
	\includegraphics[width=\textwidth, height =\textwidth]{images/stacked_GW_plot_1000}
	\caption{As Figure \ref{fig:corner_plot_1} but for 9 realisations of the noise processes. The inferred posterior distributions show strong agreement between the different noise realisations for the majority of static parameters. The exceptions are $\iota$ and $h$ which exhibit a larger degree of variance due to the weak identifiability between these two parameters c.f. Figure \ref{fig:likelihood_surface}.} 
	\label{fig:corner_plot_2}
\end{figure*}
\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{images/likelihood_surface}
	\caption{Likelihood $\log \mathcal{L}$ (Equation \ref{eq:likelihood}) surface across the $\iota - h_0$ parameter space for a single noise realisation. The red point labels the location of the likelihood maxima. The likelihood surface has been normalised with respect to the absolute value of the maxima. A strong likelihood ridge is evident in the parameter space where the Kalman filter produces similar likelihood values, making it difficult for sampling algorithms to converge to a consistent value.}
	\label{fig:likelihood_surface}
\end{figure}
\begin{figure*}
	\centering
	\includegraphics[width=\textwidth]{images/compare_iotas_lines2}
	\caption{One-dimensional marginalized posteriors for $\iota$ over 9 realisations of the noise where $h$ is (a) an unknown free parameter (left panel) and (b) fixed at its true value $= 10^{-12}$ (right panel). The thick vertical orange line labels the true injected value. The dashed vertical lines denote the median of the corresponding coloured posterior. Fixing $h$ reduces the variance between different noise realisations but results in a clear bias between the true and inferred values of $\iota$. The left hand panel is equivalent to the marginalized posteriors for $\iota$ shown in Figure \ref{fig:corner_plot_2} (4th row, 4th column), but with the $(0.16,0.84)$ quantiles replaced with the $0.5$ quantile.}
	\label{fig:iota}
\end{figure*}
The above discussion was for a single realisation of the noise processes. It is important to confirm that our method is robust against different realisations of the noise and that the specific realisation of the noisy data used in Sections \ref{sec:parameter_estim} and \ref{sec:detection} is not particularly advantageous or convenient for our method. In this Section we confirm that we are able to infer comparable results for different noise realisations. To this end we take our representative example from Table \ref{tab:parameters_and_priors} and generate 1000 realisations of the process noise and measurement noise. For each realisation of the data we can then independently estimate the static parameters, $\boldsymbol{\theta}$. In Figure \ref{fig:corner_plot_2} we show the results for the  estimated values of $\boldsymbol{\theta}_{\rm gw}$ for nine realisations of the noise. This figure is analogous to the single noise realisation of Figure \ref{fig:corner_plot_1}. We plot only 9 realisations of the noise rather than the full set of 1000 realisations so as to not overcrowd the figure. As in Section \ref{sec:parameter_estim} the estimates of $\boldsymbol{\theta}_{\rm psr}$ are recovered unambiguously across the 1000 noise realisations and for brevity we do not show the results here. \newline

@@

We saw previously strong biases in $\iota$ and $h_0$ when considering just 9 noise realisations. These biases persist with a similar magnitude when we consider the full set of $10^3$ noise realisations. Moreover, smaller biases are also observed in $\psi$ ($\sim 0.1$ radians) and $\alpha$ ($\sim 0.04$ radians) as well as smaller still biases in $\delta$ and $\Phi_0$. These biases are a result of dropping the pulsar terms from the measurement equation and will be discussed in Section \ref{sec:bias}
\subsection{Exploring a broader parameter space} \label{sec:parameter_space}
\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{images/pp_plot_new}
	\caption{PP plot for 200 simulated GW sources for a subset of parameters of $\boldsymbol{\theta}_{\rm gw}$, randomly drawn from the prior distributions of Table \ref{tab:parameters_and_priors}. Each coloured line corresponds to a different parameter of $\boldsymbol{\theta}_{\rm gw}$. The parameters $h_0$ and $\iota$ are fixed at $5 \times 10^{-15}$ and 1.0 rad in order to maintain an approximately constant SNR. The grey shaded contours labels the $1,2,3$-$\sigma$ confidence intervals. Well estimated posteriors should fall along the diagonal $y=x$. Many parameters show evidence of being over-constrained due to a modelling bias. }
	\label{fig:parameter_space}
\end{figure}
We have so far focussed on a single representative system, summarised in Table \ref{tab:parameters_and_priors}. It is also important to test the method in different regions of the parameter space. To this end we consider 200 injections for different sets of parameter values where we fix $h_0 = 5 \times 10^{-15}$, $\iota =1.0$ and draw the remaining five parameters of $\boldsymbol{\theta}_{\rm gw}$ from the prior distributions described in Table \ref{tab:parameters_and_priors}. We fix $h_0$ and $\iota$ in order to maintain an approximately constant SNR across the parameter space. For each simulated injection we can then attempt to recover the posterior distributions for each parameter. To summarise the results across the parameter space we use a parameter-parameter (PP) plot \citep{doi:10.1198/106186006X136976}. A PP plot describes the fraction of the total number of injected parameters which are included within a given credible interval of the estimated posterior with respect to the credible interval itself. In the ideal case the PP plot should be a diagonal line, indicating that $x$-percentage of the simulated injections fall within the $x$-percentage credible interval. The results are shown in Figure \ref{fig:parameter_space}. The shaded grey contoured regions label the first 3 $\sigma$ significance levels i.e. how much deviation from the $y=x$ diagonal one might expect statistically given 200 simulations.  We can see that it only $\Omega$ is well-behaved and falls within the shaded region. The other parameters deviate from the $y=x$ diagonal. The effect is most pronounced for $\alpha$ and $\delta$ with more modest deviations for $\psi$ and $\Phi_0$. The shape of the graph indicates that the posteriors for these parameters are over-constrained; there are fewer injections contained within high value credible intervals than would be expected statistically, and vice versa. This is an result of the bias that we observed in Section \ref{sec:multiple_noise}; the posteriors inferred for these parameters are generally highly confident i.e. narrow, but are systematically biased away from the true injection values. This manifests in the PP plot as posteriors which are insufficiently narrow to capture the injected value within the appropriate credible interval. An in-depth discussion on the origin of this bias is given in Section \ref{sec:bias}.

\subsection{Bias by dropping pulsar terms}\label{sec:bias}
\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{images/likelihood_iota_new}
	\caption{Variation in the log likelihood returned by the Kalman filter using the Earth terms model (Equation \ref{eq:measuremen_earth}, blue line), the pulsar terms model (Equation \ref{eq:measurement}, orange line), and the Earth terms model for Earth terms only synthetic data (green line), with respect to the inclination $\iota$, over the prior space. All other parameters are fixed at their true injected value from Table \ref{tab:parameters_and_priors}. The coloured dashed vertical lines label the location of the likelihood maxima for their corresponding model. For the Earth-terms model the likelihood maxima does not coincide with the injected value ($\iota = 1.0$). For the pulsar-terms model the location of the maxima and injected value do coincide. 10 noise realisations are plotted but the solutions are completely overlaid.}
	\label{fig:likelihood_surface_iota}
\end{figure}
Evident in the preceding discussion is the fact that for some of the $\boldsymbol{\theta}_{\rm gw}$ parameters the inferred value (i.e. the median of the one-dimensional marginalised posterior) is biased away from the true injection value. This effect is most severe for $\iota$ and $h_0$ but also present to a lesser extent in other static parameters such as $\delta,\alpha$ and $\psi$, and additionally to an even smaller extent in $\Phi_0$. In this Section we demonstrate how this bias results from the dropping of the pulsar terms described in Section \ref{sec:parameter_estim}. \newline

In Figure \ref{fig:likelihood_surface_iota} we plot the variation in the log-likelihood returned by the Kalman filter, Equation \eqref{eq:likelihood}, as we vary $\iota$ across the prior space, holding all other parameters constant at the true injected value of the representative example system, Table \ref{tab:parameters_and_priors}. We do this for 10 separate realisations of the system noise. We consider 3 separate solutions: \textit{(i)} The solid blue line is the likelihood curve with just the Earth-terms included (i.e. using Equation \ref{eq:measuremen_earth} in the Kalman filter), \textit{(ii)} the solid orange line is the likelihood curve with the additional pulsar terms included (i.e. using Equation \ref{eq:measurement}), \textit{(iii)} the solid green line is the solution where we drop the pulsar terms from both the synthetic data and the measurement equation. Whilst we have not used the nested sampling method with either solutions \textit{(ii)} or \textit{(iii)} in this work, we are able to run the Kalman filter either using the pulsar terms in the model or else dropping them from the synthetic data consider how the resulting likelihood varies for different parameter values. The dashed lines show the location of the likelihood maxima for each of the likelihood solutions; finding the location of the maxima across all static parameters is effectively the goal of any likelihood based inference method, such as nested sampling. \newline

@@

We have taken our GW source to be monochromatic, i.e. $\Omega$ is a fixed parameter with no time dependence. Whilst this is an astronomically well-justified assumption (see discussion in Section \ref{sec:plane_gw}), and appropriate for this initial methods work, it would also be of interest to extend the state-space construction to enable the GW frequency to evolve in time. The GW source is monochromatic over the decadal PTA observation period, but the timescale set by the Earth-pulsar distance is comparable to the SMBHB evolution timescale \citep{Sesana2010}. For $\Delta f_{\rm gw} < 1 / T_{\rm obs}$ (see Equation \ref{eq:f_evolution}) the GW source frequency evolution manifests as an incoherent source of noise in the pulsar terms, whilst for  $\Delta f_{\rm gw} > 1 / T_{\rm obs}$ the pulsar terms can affect the phase coherency of the Earth terms \citep{Perrodin2018}. Careful consideration of the evolution of the GW frequency will be important as we look to extend our method to include the pulsar terms in the inference model. \newline

We have also assumed that there is only a single GW source that influences the received pulsar frequencies. However, it may be possible to resolve multiple continuous GW sources concurrently \citep{PhysRevD.85.044034}. Our method naturally extends to detecting multiple GW sources simultaneously; Equation \ref{eq:measurement} is straightforwardly modified to become a linear superposition of multiple GWs. Evaluating the performance of the method under these conditions would be a further interesting pursuit. The stochastic background itself is the incoherent summation of many individual GW sources. It therefore stands to reason that it may be possible to extend this method to also search for a stochastic GW background, providing a complementary approach to detect the stochastic background without requiring a cross-correlation of the pulsar residuals to uncover the Hellings-Downs curve. \newline

@@

%https://arxiv.org/pdf/2107.03047.pdf
\section{Conclusion}
In this paper we demonstrate a new method for the detection and parameter estimation of GWs from individual, monochromatic SMBHBs. We track the evolution of the intrinsic pulsar timing noise explicitly via a state-space method, rather than subtracting the ensemble-averaged statistics. That is, we disentangle statistically the specific time-ordered realisation of the timing noise from the GW induced modulations. This enables the inference of GW parameters conditioned on the specific observed realisation of the noisy data. We introduce the Kalman filter in order to track the intrinsic state evolution of the pulsar and combine this with a Bayesian sampling framework to estimate the posterior distributions of each parameter of the system, as well as the associated evidence (marginalised likelihood) of the model. We test our method on synthetic data. We initially focus on a single, astrophysically representative, SMBHB GW source detected with the 12.5-year NANOGrav pulsars observed over a 10 year timespan. The parameters of the model are accurately recovered and the minimal detectable strain is estimated to be $h_0 \sim 4 \times 10^{-15}$. The method is then further tested over 1000 noise realisations. Consistent MAP estimates of the parameters across all noise realisations are obtained for the majority of parameters, whilst the structural weak-identifiability of $\iota$ and $h_0$ results in a larger variance in the MAP estimates between individual noise realisations for these parameters. Exploration of a broader parameter space via 200 randomly sampled parameter vectors highlights biases in some parameters as a result of dropping the pulsar terms from the inference model. The cause of these biases is discussed. The extension of this method to include the pulsar term is a key challenge for a future work.

>>>>

\subsubsection{Multiple noise realisations} \label{sec:multiple_noise}
%\begin{figure*}
%	\includegraphics[width=\textwidth, height =\textwidth]{images/stacked_GW_plot_1000}
%	\caption{As Figure \ref{fig:corner_plot_1} but for 9 realisations of the noise processes. The inferred posterior distributions show strong agreement between the different noise realisations for the majority of static parameters. The exceptions are $\iota$ and $h$ which exhibit a larger degree of variance due to the weak identifiability between these two parameters c.f. Figure \ref{fig:likelihood_surface}.} 
%	\label{fig:corner_plot_2}
%\end{figure*}

\begin{figure*}
	%\centering % Not needed
	\begin{subfigure}[b]{0.49\textwidth}
		\includegraphics[width=\textwidth]{images/stacked_GW_plot_1000}
		\caption{Kalman filter using correct estimate of the static parameters, $\hat{\boldsymbol{\theta}} = \boldsymbol{\theta}$ }
		\label{fig:6MB_BFS}
	\end{subfigure}  
	\hfill
	\begin{subfigure}[b]{0.49\textwidth}
		\includegraphics[width=\textwidth]{images/stacked_GW_plot_1000}
		\caption{Kalman filter using incorrect estimate of the static parameters, $\hat{\boldsymbol{\theta}} \neq \boldsymbol{\theta}$}
		\label{fig:25MB_bfs}
	\end{subfigure}
	\caption{As Figure \ref{fig:corner_plot_1} but for 9 realisations of the noise processes. The inferred posterior distributions show strong agreement between the different noise realisations for the majority of static parameters. The exceptions are $\iota$ and $h$ which exhibit a larger degree of variance due to the weak identifiability between these two parameters c.f. Figure \ref{fig:likelihood_surface}.}
	\label{fig:four figures2}
\end{figure*}

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{images/likelihood_surface}
	\caption{Likelihood $\log \mathcal{L}$ (Equation \ref{eq:likelihood}) surface across the $\iota - h_0$ parameter space for a single noise realisation. The red point labels the location of the likelihood maxima. The likelihood surface has been normalised with respect to the absolute value of the maxima. A strong likelihood ridge is evident in the parameter space where the Kalman filter produces similar likelihood values, making it difficult for sampling algorithms to converge to a consistent value.}
	\label{fig:likelihood_surface}
\end{figure}
\begin{figure*}
	\centering
	\includegraphics[width=\textwidth]{images/compare_iotas_lines2}
	\caption{One-dimensional marginalized posteriors for $\iota$ over 9 realisations of the noise where $h$ is (a) an unknown free parameter (left panel) and (b) fixed at its true value $= 10^{-12}$ (right panel). The thick vertical orange line labels the true injected value. The dashed vertical lines denote the median of the corresponding coloured posterior. Fixing $h$ reduces the variance between different noise realisations but results in a clear bias between the true and inferred values of $\iota$. The left hand panel is equivalent to the marginalized posteriors for $\iota$ shown in Figure \ref{fig:corner_plot_2} (4th row, 4th column), but with the $(0.16,0.84)$ quantiles replaced with the $0.5$ quantile.}
	\label{fig:iota}
\end{figure*}
The above discussion was for a single realisation of the noise processes. It is important to confirm that our method is robust against different realisations of the noise and that the specific realisation of the noisy data used in Sections \ref{sec:parameter_estim} and \ref{sec:detection} is not particularly advantageous or convenient for our method. In this Section we confirm that we are able to infer comparable results for different noise realisations. To this end we take our representative example from Table \ref{tab:parameters_and_priors} and generate 1000 realisations of the process noise and measurement noise. For each realisation of the data we can then independently estimate the static parameters, $\boldsymbol{\theta}$. In Figure \ref{fig:corner_plot_2} we show the results for the  estimated values of $\boldsymbol{\theta}_{\rm gw}$ for nine realisations of the noise. This figure is analogous to the single noise realisation of Figure \ref{fig:corner_plot_1}. We plot only 9 realisations of the noise rather than the full set of 1000 realisations so as to not overcrowd the figure. As in Section \ref{sec:parameter_estim} the estimates of $\boldsymbol{\theta}_{\rm psr}$ are recovered unambiguously across the 1000 noise realisations and for brevity we do not show the results here. \newline

@@

We saw previously strong biases in $\iota$ and $h_0$ when considering just 9 noise realisations. These biases persist with a similar magnitude when we consider the full set of $10^3$ noise realisations. Moreover, smaller biases are also observed in $\psi$ ($\sim 0.1$ radians) and $\alpha$ ($\sim 0.04$ radians) as well as smaller still biases in $\delta$ and $\Phi_0$. These biases are a result of dropping the pulsar terms from the measurement equation and will be discussed in Section \ref{sec:bias}
\subsection{Exploring a broader parameter space} \label{sec:parameter_space}
\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{images/pp_plot_new}
	\caption{PP plot for 200 simulated GW sources for a subset of parameters of $\boldsymbol{\theta}_{\rm gw}$, randomly drawn from the prior distributions of Table \ref{tab:parameters_and_priors}. Each coloured line corresponds to a different parameter of $\boldsymbol{\theta}_{\rm gw}$. The parameters $h_0$ and $\iota$ are fixed at $5 \times 10^{-15}$ and 1.0 rad in order to maintain an approximately constant SNR. The grey shaded contours labels the $1,2,3$-$\sigma$ confidence intervals. Well estimated posteriors should fall along the diagonal $y=x$. Many parameters show evidence of being over-constrained due to a modelling bias. }
	\label{fig:parameter_space}
\end{figure}
We have so far focussed on a single representative system, summarised in Table \ref{tab:parameters_and_priors}. It is also important to test the method in different regions of the parameter space. To this end we consider 200 injections for different sets of parameter values where we fix $h_0 = 5 \times 10^{-15}$, $\iota =1.0$ and draw the remaining five parameters of $\boldsymbol{\theta}_{\rm gw}$ from the prior distributions described in Table \ref{tab:parameters_and_priors}. We fix $h_0$ and $\iota$ in order to maintain an approximately constant SNR across the parameter space. For each simulated injection we can then attempt to recover the posterior distributions for each parameter. To summarise the results across the parameter space we use a parameter-parameter (PP) plot \citep{doi:10.1198/106186006X136976}. A PP plot describes the fraction of the total number of injected parameters which are included within a given credible interval of the estimated posterior with respect to the credible interval itself. In the ideal case the PP plot should be a diagonal line, indicating that $x$-percentage of the simulated injections fall within the $x$-percentage credible interval. The results are shown in Figure \ref{fig:parameter_space}. The shaded grey contours enclose the $1\sigma$, $2\sigma$, and $3\sigma$ significance levels, given 200 injections."
We can see that only $\Omega$ falls within the $3\sigma$ shaded region. The other parameters deviate from the $y=x$ diagonal. The effect is most pronounced for $\alpha$ and $\delta$ with more modest deviations for $\psi$ and $\Phi_0$. The shape of the graph indicates that the posteriors for these parameters are over-constrained; there are fewer injections contained within high value credible intervals than would be expected statistically, and there are more injections contained within the low value credible intervals. This is an result of the bias that we observed in Section \ref{sec:multiple_noise}; the posteriors inferred for these parameters are generally highly confident, i.e. narrow, but are systematically biased away from the true injection values. This manifests in the PP plot as posteriors which are overly precise (narrow) to contain the injected value within the appropriate credible interval. An in-depth discussion on the origin of this bias is given in Section \ref{sec:bias}.

\subsection{Bias by dropping pulsar terms}\label{sec:bias}
\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{images/likelihood_iota_new}
	\caption{Variation in the log likelihood returned by the Kalman filter using the Earth terms model (Equation \ref{eq:measuremen_earth}, blue line), the pulsar terms model (Equation \ref{eq:measurement}, orange line), and the Earth terms model for Earth terms only synthetic data (green line), with respect to the inclination $\iota$, over the prior space. All other parameters are fixed at their true injected value from Table \ref{tab:parameters_and_priors}. The coloured dashed vertical lines label the location of the likelihood maxima for their corresponding model. For the Earth-terms model the likelihood maxima does not coincide with the injected value ($\iota = 1.0$). For the pulsar-terms model the location of the maxima and injected value do coincide. 10 noise realisations are plotted but the solutions are completely overlaid.}
	\label{fig:likelihood_surface_iota}
\end{figure}
Evident in the preceding discussion is the fact that for some of the $\boldsymbol{\theta}_{\rm gw}$ parameters the inferred value (e.g. the median of the one-dimensional marginalised posterior) is biased away from the true injection value. This effect is most severe for $\iota$ and $h_0$ but also present to a lesser extent in other static parameters such as $\delta,\alpha$ and $\psi$, and additionally to an even smaller extent in $\Phi_0$. In this Section we demonstrate how this bias results from the dropping of the pulsar terms described in Section \ref{sec:parameter_estim}. \newline

In Figure \ref{fig:likelihood_surface_iota} we plot the variation in the log-likelihood returned by the Kalman filter, Equation \eqref{eq:likelihood}, as we vary $\iota$ across the prior space, holding all other parameters constant at the true injected value of the representative example system, Table \ref{tab:parameters_and_priors}. We do this for 10 separate realisations of the system noise. We consider three separate situations: \textit{(i)} The solid blue line is the likelihood curve with just the Earth-terms included (i.e. using Equation \ref{eq:measuremen_earth} in the Kalman filter), \textit{(ii)} the solid orange line is the likelihood curve with the additional pulsar terms included (i.e. using Equation \ref{eq:measurement}), \textit{(iii)} the solid green line is the solution where we drop the pulsar terms from both the synthetic data and the measurement equation. Whilst we have not used the nested sampling method with either solutions \textit{(ii)} or \textit{(iii)} in this work, we are able to run the Kalman filter either using the pulsar terms in the model or else dropping them from the synthetic data consider how the resulting likelihood varies for different parameter values. The dashed lines show the location of the likelihood maxima for each of the likelihood solutions; finding the location of the maxima across all static parameters is effectively the goal of any likelihood based inference method, such as nested sampling. \newline

@@

We have taken our GW source to be monochromatic, i.e. $\Omega$ is a fixed parameter with no time dependence. Whilst this is an astronomically well-justified assumption (see discussion in Section \ref{sec:plane_gw}), and appropriate for this initial methods work, it would also be of interest to extend the state-space construction to enable the GW frequency to evolve in time. The GW source is monochromatic over the decadal PTA observation period, but the timescale set by the Earth-pulsar distance is comparable to the SMBHB evolution timescale \citep{Sesana2010}. For $\Delta f_{\rm gw} < 1 / T_{\rm obs}$ (see Equation \eqref{eq:f_evolution}) the GW source frequency evolution manifests as an incoherent source of noise in the pulsar terms, whilst for  $\Delta f_{\rm gw} > 1 / T_{\rm obs}$ the pulsar terms can affect the phase coherency of the Earth terms \citep{Perrodin2018}. Careful consideration of the evolution of the GW frequency will be important as we look to extend our method to include the pulsar terms in the inference model. \newline

We have also assumed that there is only a single GW source that influences the received pulsar frequencies. However, it may be possible to resolve multiple continuous GW sources concurrently \citep{PhysRevD.85.044034}. Our method naturally extends to detecting multiple GW sources simultaneously; Equation \eqref{eq:measurement} is straightforwardly modified to become a linear superposition of multiple GWs. The stochastic background itself is the incoherent summation of many individual GW sources. It therefore stands to reason that it may be possible to extend this method to also search for a stochastic GW background, providing a complementary approach to detect the stochastic background without requiring a cross-correlation of the pulsar residuals to uncover the Hellings-Downs curve. \newline

@@

%https://arxiv.org/pdf/2107.03047.pdf
\section{Conclusion}
In this paper we demonstrate a new method for the detection and parameter estimation of GWs from individual, monochromatic SMBHBs. We track the evolution of the intrinsic pulsar timing noise explicitly via a state-space method, rather than subtracting the ensemble-averaged statistics. That is, we disentangle statistically the specific time-ordered realisation of the timing noise from the GW induced modulations. This enables the inference of GW parameters conditioned on the specific observed realisation of the noisy data. We introduce the Kalman filter in order to track the intrinsic state evolution of the pulsar and combine this with a Bayesian sampling framework to estimate the posterior distributions of each parameter of the system, as well as the associated evidence (marginalised likelihood) of the model. We test our method on synthetic data. We initially focus on a single, astrophysically representative, SMBHB GW source detected with the 12.5-year NANOGrav pulsars observed over a 10 year timespan. The parameters of the model are accurately recovered and the minimal detectable strain is estimated to be $h_0 \sim 4 \times 10^{-15}$ for $\iota=1.0$ rad. The method is then further tested over 1000 noise realisations. Consistent MAP estimates of the parameters across all noise realisations are obtained for the majority of parameters, whilst the structural weak-identifiability of $\iota$ and $h_0$ results in a larger variance in the MAP estimates between individual noise realisations for these parameters. Exploration of a broader parameter space via 200 randomly sampled parameter vectors highlights biases in some parameters as a result of dropping the pulsar terms from the inference model. The cause of these biases is discussed. The extension of this method to include the pulsar term is a key challenge for a future work.
